### 2021-01-16 docs/manuscript/manuscript.Rmd
writing

---
title: "A Simple Ratio for Estimating the Effective Reproduction Value of an Epidemic"
authors:
- name: Authors
  email: emails
  address: institutions
abstract: |
  ABSTRACT
acknowledgements: |
  Acknowledgements
output:
  bookdown::pdf_book:
    base_format: rticles::oup_article
    number_sections: no
    extra_dependencies: float
    keep_tex: yes
    includes:
      in_header: header.tex
  bookdown::word_document2:
    number_sections: no
    extra_dependencies: float
    keep_tex: yes
    includes:
      in_header: header.tex
header-includes:
  - \usepackage[labelsep=period,labelfont=bf]{caption}
  - \usepackage{hyperref}
  - \usepackage[capitalize, nameinlink]{cleveref}
  - \crefdefaultlabelformat{#2\textbf{#1}#3} % <-- Only #1 in \textbf
  - \crefname{figure}{\textbf{Figure}}{\textbf{Figures}}
  - \Crefname{figure}{\textbf{Figure}}{\textbf{Figures}}
  - \crefname{table}{\textbf{Table}}{\textbf{Tables}}
  - \Crefname{table}{\textbf{Table}}{\textbf{Tables}}
bibliography:
  - packages.bib
  - references.bib
keywords:
  - key
  - dictionary
  - word
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # By default, hide code; set to TRUE to see code
knitr::opts_chunk$set(fig.pos = 'p') # Places figures on pages separate from text
knitr::opts_chunk$set(out.width = '100%', dpi=300) # Figure resolution and size
knitr::opts_chunk$set(fig.env="figure") # Latex figure environment
knitr::opts_chunk$set(eval.after = "fig.cap")
knitr::opts_chunk$set(echo = F,
                      message = F, 
                      warning = F)

## Load packages
library("tidyverse")
library("ggplot2")
library("ggpubr")
library("drake")
library("english")
library("bibtex")
library("here")

# Set working directory
here::i_am("docs/manuscript/manuscript.Rmd")

# Get Figure creation functions
source(here::here("code", "create_figures.R"))

# Convert number to text 
# Ref: https://5harad.com/mse125/r/visualization_code.html
addUnits <- function(n) {
  labels <- ifelse(n < 1000, n,  # less than thousands
                   ifelse(
                     n < 1e6,
                     paste0(round(n / 1e3), ' thousand'),
                     # in thousands
                     ifelse(
                       n < 1e9,
                       paste0(round(n / 1e6), ' million'),
                       # in millions
                       ifelse(n < 1e12, paste0(round(n / 1e9), ' billion'), # in billions
                              'too big!')
                     )
                   ))
  return(labels)
}

# Covert a vector to a english list
vector_to_englishList <- function(x) {
  last_item = tail(x, 1)

  englishList = paste(head(x,-1), collapse = ", ")
  englishList = paste(englishList, last_item, sep = ", and ")

  return(englishList)
}

# Function between to numbers
```

```{r load-data}
simulation_parameters = readd("simulation_parameters")
rt_estimates = readd("rt_estimates")
national_data = readd("national_data")
COVID19_params = readd("COVID19_params")

# Comparison with COVID-19 data
national_data_filtered = national_data %>% 
  filter(time > as.Date("2020-04-01") & time < as.Date("2020-12-01")) %>% 
  filter(cases_new > 19)

estimation_diff = national_data_filtered$estimation_diff
estimation_diff_median = estimation_diff %>%  median(., na.rm=T) %>% signif(digits = 2)
estimation_diff_quantile_975 = estimation_diff %>% quantile(., 0.975, na.rm = T) %>% signif(digits = 2) %>% format(scientific = F)
estimation_diff_quantile_025 = estimation_diff %>% quantile(., 0.025, na.rm = T) %>% signif(digits = 2) %>% format(scientific = F)

# Get packages used for this project
packages <- readLines(here::here("code", "packages.R")) %>% #read file
              sub('[^\"]+\"([^\"]+).*', '\\1', .)    
package_citation_ids =  read.bib(here("./docs/manuscript", "packages.bib")) %>% 
  names(.) %>% regmatches(., m = regexpr(pattern = "^R-.+", text = .)) %>% 
  paste("@", ., sep="") %>% paste(collapse = "; ")
```

# Introduction {.unnumbered}

The effective reproduction number ($R_t$) is a fundamental measure in infectious disease epidemiology. It can be defined as the average number of secondary cases caused by an infected individual. The $R_t$ changes over the course of an epidemic, because it is not only determined by characteristics of the pathogen but also by human behavior (e.g. face mask usage, physical distancing, and travel restrictions). If the $R_t$ is above 1, it indicates that the epidemic is growing exponentially. If, by contrast, the $R_t$ drop below 1, it indicates that the number of new cases is decreasing. Furthermore, the $R_t$ can vary between geographic locations, as social behaviors and environmental factors vary between these locations.

From a public health perspective, the $R_t$ can be a method to communicate whether an epidemic is growing in a region. However, this communication may be hindered by the fact that many of the current methods for estimating $R_t$ require advanced mathematics to understand and to apply [@cori_new_2013; @wallinga_different_2004; @wallinga_how_2007]. Typically these methods require a description of the generation interval distribution, which is the distribution of times between when a primary case becomes infectious to when a secondary case becomes infectious.

Here, we describe how $R_t$ can be approximated by a simple ratio of new infections on day $t$ and day $t-d$, where $d$ is the mean generation interval of the infection (\cref{eq:simple-ratio-eq}). Our aim is two-fold: First, by describing a simplified method of calculating $R_t$, it may allow others to easily and rapidly calculate $R_t$. Second, to assess the accuracy of the simple ratio compared to the more advanced methods for calculating $R_t$. This second aim is important, because this simplified ratio has been used to calculate $R_t$ in public health reports but we have not found significant literature describing the performance of the method (at least in English publications).

```{=tex}
\begin{equation}
  R_t \approx \frac{I_t}{I_{t-d}}
  (\#eq:simple-ratio-eq)
\end{equation}
```
The major advantage of this approximation is that its simplicity makes it easily understandable and verifiable. Anybody with basic math skills and access to the new case counts can calculate this ratio to estimate the $R_t$.

## A Simple Ratio for Estimating Rt {.unnumbered}

As described by previous publications [@wallinga_how_2007], we can define a function $n(a)$ that describes the expected number of new cases produced today (time $t$) by a single person who was infected $a$ days ago (at time $t-a$). Therefore, to calculate $I_{ta}$ (the expected number of new infections created today by all individuals who became infected $a$ days ago), we multiply $n(a)$ by the number of new infections created $a$ days ago ($I_{t-a}$).

```{=tex}
\begin{equation}
  I_{ta} = I_{t-a}n(a)
  (\#eq:eq2)
\end{equation}
```
Then to determine the total expected number of new infections today ($I_t$), we sum $I_{ta}$ for all values of $a$, as shown in \cref{eq:renewal-eq-1,eq:renewal-eq-2}. In other words, $I_t$ can be calculated by taking all individuals infected prior to $t$, calculating the expected number of new infections each one will produce based on the age of their infection, and summing all those values together.

```{=tex}
\begin{equation} 
  I_t = \sum_{a=0}^\infty{I_{ta}}da
  (\#eq:renewal-eq-1)
\end{equation}
```
```{=tex}
\begin{equation} 
  I_t = \sum_{a=0}^\infty{I_{t-a}n(a)}da
  (\#eq:renewal-eq-2)
\end{equation}
```
Next we can substitute the generation interval function ($g(a)$) into \cref{eq:renewal-eq-2} by recognizing that the generation interval distribution is equal to $n(a)$ normalized by the total number of infections a single infected person is expected cause (i.e. $R_t$).

```{=tex}
\begin{equation} 
  g(a) = \frac{n(a)}{R_t}
  (\#eq:generation-interval-eq)
\end{equation}
```
```{=tex}
\begin{equation} 
  I_t = \sum_{a=0}^\infty{I_{t-a}g(a)R_t}da
  (\#eq:sub-in-rt-1)
\end{equation}
```
```{=tex}
\begin{equation} 
  \frac{I_t}{R_t} = \sum_{a=0}^\infty{I_{t-a}g(a)}da
  (\#eq:sub-in-rt-2)
\end{equation}
```
```{=tex}
\begin{equation} 
  R_t = \frac{I_t}{\sum{I_{t-a}g(a)}da}
  (\#eq:sub-in-rt-3)
\end{equation}
```
If we make the assumption that there is minimal variance in the generation interval distribution (i.e. $g(a)$ equals approximately 1 when $a$ equals the mean generation interval ($d$) and $g(a)$ equals approximately zero for all other values of $a$), then Equation \cref{eq:sub-in-rt-3} can be simplified to:

```{=tex}
\begin{equation}
  R_t \approx \frac{I_t}{I_{t-d}}
  (\#eq:simple-ratio-eq)
\end{equation}
```
We can make this estimation of $R_t$ robust to variations in the generation interval and reporting of new cases by calculating $R_t$ as the ratio of the sums of new cases within two time windows size $\tau$ days (with the center of the time windows being on days $t$ and $t-d$):

```{=tex}
\begin{equation} 
  R_t \approx \frac{\sum_{t - \frac{\tau}{2}}^{t + \frac{\tau}{2}}I_i}{\sum_{t - \frac{\tau}{2}}^{t + \frac{\tau}{2}}I_{i-d}} = \frac{n_{t}}{n_{t-d}}
  (\#eq:eq6)
\end{equation}
```
A confidence interval can be generated for this ratio by carrying forward a confidence interval in the generation interval.

# Materials and methods {.unnumbered}

## Simulation {.unnumbered}

A deterministic Susceptible-Exposed-Infectious-Resolved (SEIR) model was used to create epidemic simulations. No birth rate, death rate, or R to S transition rates were included. We included `r simulation_parameters$Nsims` simulated epidemics in our analysis. For each epidemic, the mean generation interval of the infection was randomly selected from `r simulation_parameters$mean_gi_range[1]` to `r simulation_parameters$mean_gi_range[2]` days. The generation interval was randomly divided into latent and infectious periods. Each simulation was divided in to `r as.english(simulation_parameters$num_periods[1])` `r simulation_parameters$period_lengths[1]` day periods with differing $R_0$ values (representing changes in social behaviors in the populations); and the $R_0$ in these periods was randomly selected from `r simulation_parameters$r0_range[1]` to `r simulation_parameters$r0_range[2]`. Between each `r simulation_parameters$period_lengths[1]` day period, a transition period was included in which the $R_0$ would gradually change following a logistic curve. The length of the transition period was randomly selected from `r simulation_parameters$transition_lengths[1]` to `r simulation_parameters$transition_lengths[2]` days. Only simulations that lasted at least `r simulation_parameters$min_simLength` days and never had less than `r simulation_parameters$min_newCases` cases per day were included in final analysis.

## Rt Estimation {.unnumbered}

Custom R code was used to generate the simple ratio $R_t$ estimations, and the EpiEstim R package was used to generate the Cori method $R_t$ estimations. See below for information on code availability. When assessing the accuracy of both methods, estimations made prior to the 15th day of each simulation was not included because both methods, as implemented here, are significantly less accurate during this early period of the epidemic.

## COVID-19 Data {.unnumbered}

Data on daily COVID-19 cases were acquired from the covidregionaldata R package [@R-covidregionaldata]

## Code and Data Availability {.unnumbered}

All simulations, data processing, analysis, and graphics were all created using the R statistical software using the following packages: `r vector_to_englishList(packages)` [`r package_citation_ids`]. All code and data used to develop this project is available at <https://github.com/aryazand/simple_rt>

# Results {.unnumbered}

## Accuracy of the simple ratio estimation of $R_t$ {.unnumbered}

```{r}
# Descriptive Statistics
N = simulation_parameters$initial_state_values %>% sum()
error_median = with(rt_estimates, median(abs(rt.simple_mean_error), na.rm = T)) %>% signif(digits = 2)
error_quantile_975 = with(rt_estimates, quantile(abs(rt.simple_mean_error), 0.975, na.rm = T)) %>% signif(digits = 2) %>% format(scientific = F)
error_quantile_025 = with(rt_estimates, quantile(abs(rt.simple_mean_error), 0.025, na.rm = T)) %>% signif(digits = 2) %>% format(scientific = F)
```

We assessed how accurately the simple ratio can estimate $R_t$ compared to a previously developed method by Cori et al [-@cori_new_2013], which is a popular method in the field - we will refer to this method as the "Cori method". $R_t$ calculation with either the simple ratio and Cori method requires knowing the mean generation interval of the infection; however, the Cori method also requires knowing the standard deviation of the generation interval or having an empiric distribution of the generation interval.

To assess accuracy of $R_t$ estimations, we used Susceptible-Exposed-Infectious-Resolved (SEIR) model to simulate epidemics and then compare how accurately the two methods can estimate the $R_t$ values in the epidemics. We simulated `r simulation_parameters$Nsims` epidemics in a population of `r addUnits(N)` individuals. We varied the lengths of latent and infectiousness periods with each simulation. In each simulation, the population underwent changes in the $R_t$ value representing changes in population behavior (e.g. "lockdown" period or lifting of a lockdown). \cref{fig:sim-results}**A** shows the true $R_t$ and the estimated $R_t$ values for each day in one such simulated epidemic. A full visualization of the all the simulations can be seen in (\cref{fig:all-simulations}).

Over the `r simulation_parameters$Nsims` simulation, the median estimation error of the daily $R_t$ by the simple ratio method was `r error_median` (90% range of `r error_quantile_025` to `r error_quantile_975`) (\cref{fig:sim-results}**B**). The simple ratio has a tendency to over-estimate the $R_t$, especially as the true value of $R_t$ increases above 1.5 (\cref{fig:sim-results}**C**). This is consistent with previous mathematical analysis of this simplified method of estimating $R_t$ [@wallinga_how_2007].

(ref:sim-results-cap) \textbf{The Simple Ratio approximates the effective reproduction in simualted epidemics.} One hundred epidemics were simulated using an SEIR model and the simple ratio and Cori method were used to estimate the known $R_t$ over time for each simulated epidemic. \textbf{(A)} The true $R_t$ (black line) and estimated $R_t$ (colored lines) over time for one of simulated epidemics. \textbf{(B)} The distribution of the absolute errors of both estimation methods across all simulated epidemics. \textbf{(C)} The estimation error of the simple ratio method is associated with size of true $R_t$

```{r sim-results, fig.cap = '(ref:sim-results-cap)'}
# Plots
panel_A <- rt_estimates %>%
  filter(simulation_run == 66) %>%
  create_estimate_plot()

panel_B <- rt_estimates %>%
  filter(time > 14) %>%
  create_rt_estimation_error_boxplot()

panel_C <- rt_estimates %>%
  filter(time > 14) %>%
  create_rt_estimation_error_by_rt()

panel_D <- rt_estimates %>% 
  pivot_longer(cols = ends_with("mean_error"), names_to = "Method", values_to = "estimation_error") %>% 
  mutate(Method = recode(Method, rt.cori_mean_error = "cori", rt.simple_mean_error = "simple")) %>%
  ggplot(aes(x = time, y = estimation_error)) + geom_line(aes(group = interaction(simulation_run, Method), color = Method), alpha = 0.2) + 
  facet_wrap(~Method, ncol = 2) +
  ylab("Estimation Error") + 
  custom_theme() + 
  theme(strip.background = element_blank(), 
        strip.text = element_blank(),
        legend.position = "none") + 
  scale_color_manual(values=cbPalette) +
  scale_fill_manual(values=cbPalette) + 
  ylim(-0.25, 0.5)

full_figure_top = ggarrange(panel_A, panel_B, panel_C, ncol = 3, align = "hv", labels=c("A", "B", "C"), common.legend = T)
full_figure_bottom = ggarrange(panel_D, labels=c("D"))
ggarrange(full_figure_top, full_figure_bottom, nrow = 2, common.legend = T)

```

Often however, the the mean and standard deviation of the generation interval is typically not known exactly. For instance, for the COVID-19 the mean generation interval estimates, with 95% credible intervals, include 3.95 (3.01 - 4.91) and 5.20 (3.78 - 6.78); and estimates of the standard deviation of the generation interval include 1.51 (0.74 - 2.97) and 1.72 (0.91 - 3.93) [@ganyani_estimating_2020]. Therefore, we compared the accuracy of the two methods if the standard deviation of the generation interval is mis-specified as three days longer. Under these conditions, the simple ratio actually performs slightly better than than the Cori method (\cref{fig:estimation-inaccParams}).

```{r real-world-data, fig.cap = "\\textbf{The Simple Ratio and Cori Method estimate similar Rt values for the COVID-19 pandemic}. \\textbf{(A)} Rt esimations for four sample nations \\textbf{(B)} Histogram displaying the absolute difference in Rt esimation across for each day of the pandemic (Apr 01 to Dec 12) for all 233 individual nations."}

countries = c("United States", "Canada", "United Kingdom", "Germany")

panel_A <- national_data %>%
  filter(country %in% countries) %>%
  filter(time %in% seq(as.Date("2020-10-01"), as.Date("2020-10-31"), by = "day")) %>%
  pivot_longer(., cols = starts_with("rt."), names_to = c("Method", "parameter"), 
              names_pattern = "rt.([[:alpha:]]+)_([[:print:]]+)", values_to = "value") %>%
  pivot_wider(names_from = "parameter", values_from = "value") %>%
  ggplot() + 
    geom_line(aes(time, mean, group = Method, color = Method), size = 1) +
    geom_ribbon(aes(time, ymin = quantile_025, ymax = quantile_975, group = Method, fill = Method), alpha = 0.2) + 
    geom_hline(yintercept = 1, linetype = 2) + 
    facet_wrap(~country) + 
    ylab("Rt") + 
    custom_theme() + 
    theme(strip.background = element_blank(),
          legend.position = "top") +
    scale_color_manual(values=cbPalette)

panel_B <- national_data_filtered %>%
  ggplot() + 
    geom_histogram(aes(estimation_diff)) + scale_x_log10(labels = scales::label_number(accuracy = 0.0001)) + 
    labs(x = "absolute difference in Rt estimation") +
    custom_theme()

ggarrange(panel_A, ggarrange(panel_B, ncol=2), nrow = 2, heights = c(2,1), labels = c("A","B"))
```

## Using the Simple Ratio to Estimate $R_t$ with real-world COVID-19 data {.unnumbered}

We next compared $R_t$ estimates from the Simple Ratio and Cori et al for using data from the current COVID-19 pandemic. We estimated $R_t$ in four nations using a generation interval with mean of `r COVID19_params$mean_gi` and standard deviation of `r COVID19_params$sd_gi`[@ganyani_estimating_2020]. \cref{fig:real-world-data}**A** shows the results for the month of October 2020. Visual inspection shows that the two methods of estimating $R_t$ have very good concordance. Across all nations over the course of the pandemic, the median difference in estimation between the two methods was `r estimation_diff_median` (95% range of `r estimation_diff_quantile_025` to `r estimation_diff_quantile_975`) (\cref{fig:real-world-data}**B**).

# Discussion {.unnumbered}

# References {.unnumbered}

::: {#refs}
:::

# Supplementary Material {.unnumbered}

\beginsupplement

```{=tex}
\newpage
\blandscape
```
```{r all-simulations, fig.width = 11, fig.height = 6, fig.cap = "All Simulations"}

rt_estimates %>%
  create_estimate_plot() + 
  facet_wrap(~simulation_run) + 
  theme(strip.background = element_blank(), strip.text = element_blank())
```

\elandscape

\newpage

```{r estimation-inaccParams,  fig.cap = "(A) estimation with gi mean mis-specified. (B) estimation with gi sd mis-specified."}

rt_estimates_meanerr <- readd("rt_estimates_meanerr")
rt_estimates_sderr <- readd("rt_estimates_sderr")

panel_A <- rt_estimates_meanerr %>% 
  create_rt_estimation_error_boxplot()

panel_B <- rt_estimates_sderr %>% 
  create_rt_estimation_error_boxplot()

ggarrange(panel_A, panel_B, ncol=2, align = "hv", labels=c("A", "B"), common.legend = T)
```